---
title: Likelihood Profile to Generate Standard Error
author: "Metrum Research Group"
date: ""
---


```{r, setup, echo = FALSE, message = FALSE}
source("script/global.R")
knitr::opts_chunk$set(fig.path = "figures/intro-")
```



```{r}
library(tidyverse)
library(mrgsolve)
library(nloptr)
options(mrgsolve.soloc = "build")
```

# First, we fit the model by normal likelihood estimation

```{r}
mod <- mread_cache("model/yoshikado.cpp", end = 12, delta = 0.5)
data <- read_csv("data/fig4a.csv") 
data <- mutate(data, DV = ifelse(DV < 0, NA_real_, DV), profile = NULL)
dose <- filter(data, evid==1)

fun <- function(p,data,mod,name) {
  names(p) <- name
  p <- lapply(p,exp)
  mod <- param(mod, p)
  out <- mrgsim_d(mod,data,output="df")
  like <- dnorm(log(data[["DV"]]),log(out[["CP"]]),sd = p$sigma,log=TRUE)
  -2*sum(like, na.rm=TRUE)
}

theta <- c(
  fbCLintall = 1.2, 
  ikiu = 1.2, 
  fbile = 0.9, 
  ka = 0.1, 
  ktr = 0.1,
  sigma = 1
) %>% log()

fun(theta,data,mod,names(theta))

fit <- newuoa(x0 = theta, fn = fun, data = data, mod = mod, name=names(theta))
```


# Generate CI by Hessian for Comparison
Get a 95% CI using Hessian by finite difference
```{r}
parn <- 2

mod2 <- update(mod, rtol = 1E-14, atol = 1e-14, hmax = 0.01)

h <- nlme::fdHess(fit$par, fun = fun, data = data, mod = mod2, name = names(theta))

se <- h$Hessian %>% solve %>% diag %>% sqrt 

ci <- fit$par[parn] + 1.96*se[parn]*c(-1,1)

names(ci) <- c("lb", "ub")

exp(ci)
```

Same answer with `numDeriv`

```{r}
h2 <- numDeriv::hessian(fun, fit$par, data = data, mod = mod2, name = names(theta))
se2 <- h2 %>% solve %>% diag %>% sqrt
ci2 <- fit$par[parn] + 1.96*se2[parn]*c(-1,1)
exp(ci2)
```


# Generate CI by likelihood profile


First, let's take the estimates from the maximum likelihood estimation
as the new set of initial estimates. 

But we want to remove `ikiu`

```{r}
th <- fit$par
names(th) <- names(theta)
th <- th[names(th) != "ikiu"]
```

Pick a bunch of values for `ikiu` for evaluating the likelihood

```{r}
ikiu <- exp(fit$par[parn] + seq(-1,1,0.1))

llprof <- function(value, fix = list(ikiu = mod$ikiu)) {
  fix[[1]] <- value
  mod <- update(mod, param = fix)
  fit <- newuoa(x0 = th, fn = fun, data = data, mod = mod, name= names(th))
  tibble(par = names(fix)[1], value = value, ofv = fit$value)
}

df <- parallel::mclapply(ikiu, llprof) %>% bind_rows()

target <- 3.84 + fit$value

ggplot(df, aes(value,ofv)) + geom_line(lwd=1) + geom_point(size=4) + 
  geom_hline(yintercept = target, col="firebrick") + 
  geom_vline(xintercept=exp(fit$par[parn]),col="firebrick") 

```

```{r}
lo <- filter(df, value < mod$ikiu)
hi <- filter(df, value > mod$ikiu)

lb <- approx(lo[["ofv"]], lo[["value"]], target)$y
ub <- approx(hi[["ofv"]], hi[["value"]], target)$y


list(hess = exp(ci), prof = c(lower = lb, uppper = ub))
```

# Compare
```{r}
ggplot(df, aes(value,ofv)) + geom_line(lwd=1) + geom_point(size=4) + 
  geom_hline(yintercept = target, col="firebrick") + 
  geom_vline(xintercept = exp(fit$par[parn]),col="firebrick") + 
  geom_vline(xintercept = exp(ci),col="darkgreen",lwd=1) +
  geom_vline(xintercept = exp(ci2),col="darkgrey",lty=2,  lwd=0.8) +
  geom_vline(xintercept = c(lb,ub),col="orange",lwd=1) +
  ggtitle("Green/Grey: hessian; Orange: likelihood profile")
```

